%========================================================================================
\section{Supporting Information on Chapter 1} \label{Annex_chap1}
%========================================================================================

%----------------------------------------------------------------------------------------
%\subsection*{Aflatoxin levels in food commodities} 
%----------------------------------------------------------------------------------------

%++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++
\input{tables/aflatoxin_food_levels} 
%++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++

%----------------------------------------------------------------------------------------
\subsection*{National and Internationally Harmonized Limits for Aflatoxins in Foodstuffs} % Main appendix title
%----------------------------------------------------------------------------------------

To gain an overview of the changes in the legal limits for aflatoxins in food, both at the national and international level, over a period of 20 years, a comparative analysis was carried out between the years 2002 and 2022. The data research revealed that in most countries, regulatory limits have been established for the sum of the four major aflatoxins (AFB1, AFB2, AFG1, and AFG2) and/or for the most toxic aflatoxin (AFB1), particularly in the context of maize and peanuts. Consequently, data were collected for these specific foods and parameters. In 2002, the Food and Agriculture Organization of the United Nations (FAO) made an important contribution in this area by conducting a comprehensive study aimed at assessing the global landscape of mycotoxin regulation \citep{van2004worldwide}. This study found that for corn and peanuts, a total of 89 countries have set limits. Of these, 67 countries set national standards, while 22 countries adopted internationally harmonized standards within economic unions such as the European Union (EU), Common Market of the South (Mercosur) and Australia/New Zealand. In the following years, more countries set their own national standards, and more nations joined these economic unions, e.g. through the eastward expansion of the EU. In addition, the emergence of economic unions such as ARSO (African Organization for Standardization), EAC (East African Community), EACU (Eurasian Customs Union) and GSO (Gulf Cooperation Council Standardization Organization) contributed to the implementation of limits by more countries. The following table shows the status of regulation at the national and international level for different countries and time periods. This data was then used for creating world maps (Figure \ref{fig:Aflatoxin_Regulation_Limits}). In cases where multiple limits existed for a given time period and nation, the most stringent value was used for visualization.

\clearpage

%++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++
\input{tables/aflatoxin_regulation}
%++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++

\clearpage

%----------------------------------------------------------------------------------------
\subsection*{Physical-Chemical Property Estimation for Aflatoxins} 
%----------------------------------------------------------------------------------------
\label{Annex:estimation}

Only a limited number of experimental studies have been conducted on the physicochemical properties of AFs, necessitating the use of estimation software applications. The property estimation softwares OCHEM, EPISuite, ACD/Labs and OPERA were used to predict the boiling point (T\textsubscript{b}), melting point (T\textsubscript{m}), vapor pressure  (Log(P\textsubscript{v})), water solubility (Log(c\textsubscript{max,w})), Henry coefficient (Log(K\textsubscript{H})), octanol-air partitioning coefficient (Log(K\textsubscript{OA})), octanol-water partitioning coefficient (Log(K\textsubscript{OW})),   soil absorption coefficient (Log(K\textsubscript{OC})) of the four primary AFs (AFB1, AFB2, AFG1, and AFG2) as well as two major metabolites (AFB2a and AFM1). These estimations were either taken from freely available online databases such as OCHEM, CompTox and ChemSpider or manually calculated using the EPI Suite program (version 4.11). The estimations derived from the individual models are presented in table \ref{table:Aflatoxin_estimations}. However, using these models resulted in a high variability in the predicted values from the different calculators. In this regard, \citet{tebes2018demonstration} demonstrated for different chemicals that no individual estimation model outperforms the others, because the performance of the calculators is based on chemical class and the property value. However, the authors found that the geometric mean  and the median of the calculated values from these multiple calculators that use different estimation algorithms are recommended as more reliable estimates of the property value than the value from any single calculator. For that reason, the median values were used to derive the entry pathways and fate of AFs in soil as described in chapter \ref{subchap:aflatoxins_in_soil} and presented in table \ref{table:Aflatoxin_properties}.

%++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++
\input{tables/aflatoxin_estimations} 
%++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++

\clearpage
