\begin{acknowledgements}
\addchaptertocentry{\acknowledgementname} % Add the acknowledgements to the table of contents
First and foremost, I would like to express my deepest gratitude to my advisor and mentor, Katherine Muñoz, for her unparalleled support and guidance during the difficult journey to complete my dissertation. I have always greatly appreciated and admired your exceptional and inspiring way of thinking, organizational skills, and tireless diligence. Our open exchange of ideas has always spurred me to new heights and inspired my thinking. You taught me to think, work and write scientifically. I am very grateful for your unwavering belief in my abilities and your willingness to let me work independently on my projects. Working with you has been a truly valuable experience, and I greatly appreciate the support you have given me during my doctoral studies. I extend my gratitude to Prof. Dr. Gabriele Ellen Schaumann, who co-supervised me and provided me with invaluable and uplifting feedback on my endeavors.


My thanks go to my my colleagues of our working group ‘Environmental and Soil Chemistry’, I thank you for your warm welcome, cooperativeness and support during the last years. My special thanks go to Camilla More, who has faithfully stood by my side throughout my doctoral studies. From the beginning of my doctoral studies, I had the honor and pleasure of supervising your bachelor thesis - a task that eventually made us not only valued colleagues, but also good friends. Over time, your role transformed from a student assistant during your bachelor program to an indispensable technical assistant and valued colleague. My special thanks to Angelika Holderle, who helped me through all administrative and organizational issues I had all theses years and for her open, gentle and helpful nature. I would also like to thank Zacharias Steinmetz, Kilian Kenngott, Christian Buchman, Maximiliam Meyer, and my other colleagues who have contributed to an invigorating and enjoyable work environment. Their constant support, insightful discussions, and camaraderie are truly appreciated. Furthermore, I want to thank our technical assistants, Karin Meyer, Silvia Eichhöfer, Anna Baskal and Andreas Hirsch for their support, skills and professional commitment in the laboratory that helped me many times during my experiments.


I extend my gratitude to all my student assistants and past bachelor's and master's students. A special mention goes to Sven Korz, Niklaus Dahlke, Kay Seufferheld, Sarah Morlock, Selina Morscheid, Maureen Roth, and Lisa Strotkötter. Their unwavering dedication and commendable efforts not only motivated me but also prompted me to revisit and refine my concepts. They consistently challenged me to elucidate intricate subjects in a comprehensible manner. Their contributions held significance, indirectly influencing my doctoral journey by providing crucial preliminary insights that laid the groundwork for planning and executing my main experiments. I value the commitment you demonstrated under my supervision, and it was a pleasure to work alongside you.


I also extend my gratitude to the collaborators of the AflaZ project, with special recognition to Markus Schmidt-Heydt and Christian Roder. Their dedication to managing and coordinating the project played a pivotal role in its inception and successful conclusion. Your positive and friendly demeanor greatly enriched my experience in the project. Furthermore, I'd like to express my appreciation to my project colleagues Christine Schwake-Anduschus, Wolfgang Büchs, Ronald Maul, and Torsten Meiners, as well as my fellow doctoral candidates, Alexandra Schamann, Ginson Riungu, and Beatrice Tenge. Our collaborative environment fostered by numerous meetings, conferences, and trips to Kenya created an enjoyable atmosphere. Lastly, special thanks go to our Kenyan project collaborators, particularly Charles Nkonge, Hillary Kibet Cheruiyot, and Ruth Amata. Your assistance and on-site support were essential, especially during the soil sampling in Kenya when travel restrictions due to the COVID-19 pandemic prevented my direct involvement. Your contributions were crucial in bringing this project to a successful conclusion.


Abschließend möchte ich meiner Familie, meinen Freunden und insbesondere meiner Freundin Lena von ganzem Herzen für ihre unermüdliche Geduld, Unterstützung und ständige Ermutigung während meiner Promotion danken. Eure Anwesenheit hat nicht nur meine Freude vervielfacht, sondern auch meine Entschlossenheit gestärkt. In den schwierigsten Momenten gaben mir die unterstützenden Gespräche und euer anhaltender Glaube an mich Kraft. Auch die ständigen Nachfragen über die Fortschritte meiner Dissertation haben mich motiviert und mich daran erinnert, dass ich auf diesem Weg nicht allein bin. Ohne euch wäre diese Arbeit in dieser Form nicht zustande gekommen. Ich bin froh, dass es euch gibt!


\end{acknowledgements}