%----------------------------------------------------------------------------------------
\chapter*{Parts of this Thesis, Author Contributions and Third Party Assistance}
%----------------------------------------------------------------------------------------

This cumulative dissertation includes six chapters of which parts have been published as peer-reviewed research papers. In detail, Chapters 2, 4 and 5 are based on the respective accepted manuscripts, retaining their original formatting without any subsequent adjustments made. All research was conducted at the iES Landau, Institute for Environmental Sciences, at the University of Kaiserslautern--Landau (German: Rheinland--Pfälzische Technische Universität Kaiserslautern--Landau, also known as RPTU) in collaboration with the Kenya Agricultural and Livestock Research Organization (KALRO) for field sampling activities in Kenya. Financial support was provided by the "AflaZ: Zero Aflatoxin" project funded by the Federal Ministry of Food and Agriculture (BLE) under the reference AflaZ 2816PROC14. The subsequent section states the use of third party assistance and outlines my and the co-authors' specific contributions to the scientific work.

%----------------------------------------------------------------------------------------
\section*{Third Party assistance}
%----------------------------------------------------------------------------------------

In this thesis, the AI tools ChatGPT (version 3.5), Grammarly and DeepL (Translator) were utilized in the writing process. Throughout the writing process, DeepL was used for translation and Grammarly for grammar and spell checking. The use of ChatGPT was confined to the Introduction (Chapter 1), the Kenyan field study (Chapter 3), and synthesis (Chapter 6) and was not involved in Chapters that were published in peer-reviewed journals, i.e., Chapters 2, 4, and 5. Its practical applications included serving as a tool for improving the overall quality of the text. Specifically, ChatGPT was utilized for tasks such as providing suggestions for phrasing and sentence structure, finding synonyms, acting as grammar and spell-checker, and aiding in the concise summarization of content. ChatGPT supported the programming and formatting of LaTeX for the preparation of the thesis draft, especially for table formatting and the generation of BibTeX entries for the bibliography.

All figures in this dissertation were either created by the author or obtained from external sources with Creative Commons Zero (CC0) licenses, unless otherwise specified in the captions. The CC0-licensed clipart and photos were sourced from freely available online databases, including Pexels, publicdomainvectors.org, Free-images.com, and Clker.com.

%----------------------------------------------------------------------------------------
\section*{Publications}
%----------------------------------------------------------------------------------------

The following chapters of this thesis have been published in peer-reviewed journals:

{\parindent2cm \hangindent\parindent\noindent\hbox to\parindent{\textbf{\textbf{Chapter 2}}\hfil}Albert, J., More, C. A., Dahlke, N. R. P., Steinmetz, Z., Schaumann, G. E., and Muñoz, K. (2021). “Validation of a Simple and Reliable Method for the Determination of Aflatoxins in Soil and Food Matrices”. \textit{ACS Omega} 6.29, pp. 18684–18693. \texttt{DOI:} \href{https://pubs.acs.org/doi/10.1021/acsomega.1c01451}{\texttt{10.1021/acsomega.1c01451}}. \nocite{albert2021validation} \\[5pt]
\textit{Author contributions:}  J. Albert and K. Muñoz contributed equally to conceptualizing the study and designing the experiments. The experiments were conducted by J. Albert, C. A. More, and N. R. P. Dahlke. J. Albert analyzed and visualized the data. J. Albert took the lead in writing and finalizing the manuscript. Project supervision was carried out by K. Muñoz and G. E. Schaumann. K. Muñoz was responsible for project administration and funding. The manuscript was finalized through the contributions of all authors. \par} 


{\parindent2cm \hangindent\parindent\noindent\hbox to\parindent{\textbf{\textbf{Chapter 4}}\hfil}Albert, J. and Muñoz, K. (2022). “Kinetics of microbial and photochemical degradation of aflatoxin B1 in a sandy loam and clay soil”. \textit{Scientific Reports} 12.1. \texttt{DOI:} \href{https://www.nature.com/articles/s41598-022-20727-1}{\texttt{10.1038/s41598-022-20727-1}}. \nocite{albert2022kinetics}\\[5pt]
\textit{Author contributions:}  J.A. designed the study, conceived the experiments, conducted data analysis, wrote and revised the manuscript; K.M. was involved in the study design, acquired funding, supervised the project and revised the manuscript. \par}

\clearpage
 

{\parindent2cm \hangindent\parindent\noindent\hbox to\parindent{\textbf{\textbf{Chapter 5}}\hfil}Albert, J., More, C., Korz, S., and Muñoz, K. (2023). “Soil Microbial Responses to Aflatoxin Exposure: Consequences for Biomass, Activity and Catabolic Functionality”. \textit{Soil Systems} 7.1, p. 23. \texttt{DOI:} \href{https://www.mdpi.com/2571-8789/7/1/23}{\texttt{10.3390/soilsystems7010023}}. \nocite{albert2023soil}\\[5pt]
\textit{Author contributions:} J. Albert and K. Muñoz were involved in the conceptualization of the study. Experiments were carried out by J. Albert, C. A. More, and S. Korz. Data analysis and visualization were carried out by J. Albert. J. Albert took the lead in writing and finalizing the manuscript. K. Muñoz was responsible for project supervision, administration and funding. The manuscript was finalized through the contributions of all authors. \par}

\hfill 

\noindent
In addition to the publications listed above that represent Chapters of my thesis, I contributed to the following publications:

{\parindent2cm \hangindent\parindent\noindent\hbox to\parindent{\textbf{\textbf{Paper}}\hfil} Kenngott, K. G. J., Albert, J., Meyer-Wolfarth, F., Schaumann, G. E., and Muñoz, K. (2022). “\textit{Fusarium} Mycotoxins in Maize Field Soils: Method Validation and Implications for Sampling Strategy”. \textit{Toxins} 14.2, p. 130. \texttt{DOI:} \href{https://www.mdpi.com/2072-6651/14/2/130}{\texttt{10.3390/toxins14020130}}.\\[5pt]
\textit{Author contributions:} Conceptualization was jointly undertaken by K. G. J. Kenngott, J. Albert, and K. Muñoz. Experiments were carried out by K. G. J. Kenngott and J. Albert. Data analysis and visualization was carried out by K. G. J. Kenngott. K. G. J. Kenngott took lead in writing and finalizing the manuscript. Project supervision was carried out by  K. Muñoz and G. E. Schaumann. K. Muñoz was responsible for project administration and funding.
\par}


{\parindent2cm \hangindent\parindent\noindent\hbox to\parindent{\textbf{\textbf{Code}}\hfil} Steinmetz, Z., Albert, J., and Kenngott, K. (2022). Envalysis: Miscellaneous Functions for Environmental Analyses. Version 0.5.4. \texttt{DOI:} \href{https://doi.org/ft9p}{\texttt{ft9p}}. \nocite{steinmetz2021envalysis}\\[5pt] 
\textit{Author contributions:} Z. Steinmetz took the lead in setting up the package and programming and describing most of the functions. J. Albert provided code and description for the "matrix\_effect" function. J. Albert and K. G. J. Kenngott contributed to the coding and description of the "weight\_select" function. Z. Steinmetz was responsible for the maintenance of the package.
\par}
